|

### 2023-05-19 manuscript/paper.Rmd
Update Manuscript

# Models  {#sec:models}
## Imputation Model 
Let $\boldsymbol{W}$ be an $(n_{extant} + n_{fossil}) \times 1$ vector of the covariate gender of the stickleback fish and $\boldsymbol{X}$ be an $(n_{extant} + n_{fossil}) \times k$ matrix of the $k$ phenotypes of interest of the fish.  Because $\boldsymbol{W}$ contains missing data, we further define $\boldsymbol{W} = (\boldsymbol{W}_{obs}^T,\boldsymbol{W}_{mis}^T)^T$ where $\boldsymbol{W}_{obs}$ and $\boldsymbol{W}_{mis}$ are the observed and missing data of the covariate gender, respectively.  In this setting, $\boldsymbol{W}_{obs}$ is a $n_{extant} \times 1$ vector and $\boldsymbol{W}_{obs}$ is a $n_{fossil} \times 1$ vector, perfectly corresponding to the observable gender of the modern extants and the unobservable gender of the fossil data, respectively.

We impute the missing gender for the fossil data by sampling from the posterior predictive distribution $P(\boldsymbol{W}_{mis}|\boldsymbol{W}_{obs}, \boldsymbol{X})$ using the multiple imputation with chained equations (MICE) algorithm (@MICE) with predictive mean matching.  The imputation algorithm is run to obtain a total of $M = 100$ completed datasets.

@@

Let $W_{ti}$ and $X_{ti}$ be the gender and phenotype of interest for the $i$-th observation in year $t$ for $t = 1, \ldots, 18$ and  $i = 1\ldots,n_{t}$.  We assume that

>>>>

# Models  {#sec:models}
## Imputation Model 
Let $\boldsymbol{W}$ be an $(n_{extant} + n_{fossil}) \times 1$ vector of the covariate gender of the stickleback fish, $\boldsymbol{X}$ be an $(n_{extant} + n_{fossil}) \times K_{cont}$ matrix of the $K_{cont}$ continuous phenotypes of interest, and $\boldsymbol{Y}$ be an $(n_{extant} + n_{fossil}) \times K_{disc}$ matrix of the $K_{disc}$ discrete phenotypes of interest.  Because the gender of the fossilized stickleback fish is unobservable, we further define $\boldsymbol{W} = (\boldsymbol{W}_{extant}^T,\boldsymbol{W}_{fossil}^T)^T$ where $\boldsymbol{W}_{extant}$ and $\boldsymbol{W}_{fossil}$ are the $n_{extant} \times 1$ and $n_{fossil} \times 1$ vectors of the observed extant gender and missing fossil gender, respectively.

We impute the missing gender for the fossil data by sampling from the posterior predictive distribution $P(\boldsymbol{W}_{fossil}|\boldsymbol{W}_{extant}, \boldsymbol{X}, \boldsymbol{Y})$ using the multiple imputation with chained equations (MICE) algorithm (@MICE) with predictive mean matching.  The imputation algorithm is run to obtain a total of $M = 100$ completed datasets.

@@

Let $W_{ti}$, $\boldsymbol{X}_{ti}$, and $\boldsymbol{Y}_{ti}$ be the imputed gender, $K_{cont} \times 1$ vector of continuous phenotypes, and $K_{disc} \times 1$ vector of discrete phenotypes, respectively, for the $i$-th observed fossil in year $t$ for $t = 1, \ldots, 18$ and  $i = 1\ldots,n_{t}$.  In addition,
